---
title: "Assignment 1 - ESM 244"
author: "Claire Madden"
date: "1/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE)
```

## Task 1. 

snowshoe hare data : Kielland K., F. S. Chapin, R. W. Ruess. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14.  


Prepare a single polished HTML (knitted from .Rmd), planning that this might be a post/project you’d include on your personal blogdown site, that includes at least:

A useful descriptive introductory summary (3 - 4 sentences) of what’s contained in the project
Image(s), with captions and necessary citation(s), of showshoe hares and/or a map of the study area
All of your organized and well-annotated code (with warnings/messages hidden) used to create at least:

One finalized graph about the Bonanza Creek snowshoe hare population (you pick which variables, how you want to wrangle it beforehand, and which type of visual to create - but make sure it is beautifully finalized)

One finalized HTML table (probably created using kable & kableExtra) containing summary statistics about the snowshoe hares (again, you pick which variables, and how you want to group/summarize them)

Make sure that both your figure and table appear in your final knitted document, each with a useful caption. Include text associated with each to help the audience understand and interpret the results. 


```{r}

library(tidyverse)
library(kableExtra)
library(janitor)
library(lubridate)
library(tidyr)
library(ggfortify)

hares<- read_csv("showshoe_lter.csv") 

hares_mf <- hares %>% 
  filter(sex == "M" | sex == "F") %>% # remove all observations with NA in sex
  mutate(date = mdy(date)) %>% # change date format
  mutate(full_date = date) %>% # retain date column
  separate(date, into = c("year", "month", "day")) %>% #create new columns with parsed date info
  mutate(season = ifelse(month %in% c(1:4, 11, 12), "winter", "summer")) # organize months into "summer" and "winter"
  
# seems like there are a lot more observations in the summer than in the winter so maybe not an interesting thing to use for visualization



```

```{r}

# graph

hares_plot <- ggplot(data = hares_mf, aes(x = weight, y = hindft))+
  geom_point()+
  facet_wrap(~sex)

hares_plot

```







## Task 2. 

For this task, I’ll provide a couple of datasets that you *can* use for PCA exploration, but you are also welcome to find/choose a different dataset to use. You only need to use one dataset to practice running PCA, creating a biplot, and interpreting the results. Whichever dataset you choose, create a finalized HTML (knitted from .Rmd) that can be included as a project on your blogdown site including: 

A useful descriptive introductory summary (3 - 4 sentences) that helps the audience understand the data (include a citation as necessary) and what you’ll be exploring
All of your organized and well-annotated code (with warnings/messages hidden) used to wrangle then run PCA, then create a finalized PCA biplot that appears (with a caption) in your knitted HTML
A summary (bulletpoints are fine) pointing out some major takeaways from your PCA that can be gleaned from the biplot
You are also welcome to include a photo in the HTML to add some visual interest (not required for Task 2). 




```{r}
# explore data

world_env <- read_csv("world_env_vars.csv") %>% 
  clean_names(case = "upper_camel")

summary(world_env) # check out the dataset


env_subset <- world_env %>% 
  select(Elevation, TreeCanopyCover, Isothermality, RainMeanAnnual, TempMeanAnnual, Cloudiness) # select variables of interest for principal components analysis


VIM::matrixplot(env_subset, sortby = "Elevation")
# can't see variable names so not that helpful for exploration

naniar::gg_miss_var(env_subset)
# only a few NAs for each variable, will remove them

env_nona <- env_subset %>% 
  drop_na() # remove NA values



env_pca <- prcomp(env_nona, scale = TRUE) # run PCA

plot(env_pca)
biplot(env_pca) # lookin real UGLY! 

# a prettier version...
my_pca_plot <- autoplot(env_pca,
                        colour = NA,
                        loadings.colour = "black",
                        loadings.label = TRUE,
                        loadings.label.size = 3,
                        loadings.label.colour = "black",
                        loadings.label.repel = TRUE)+
  theme_minimal()
  


my_pca_plot

```





