---
title: "Assignment 1 - ESM 244"
author: "Claire Madden"
date: "1/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1. 

snowshoe hare data : Kielland K., F. S. Chapin, R. W. Ruess. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14.  


Prepare a single polished HTML (knitted from .Rmd), planning that this might be a post/project you’d include on your personal blogdown site, that includes at least:

A useful descriptive introductory summary (3 - 4 sentences) of what’s contained in the project
Image(s), with captions and necessary citation(s), of showshoe hares and/or a map of the study area
All of your organized and well-annotated code (with warnings/messages hidden) used to create at least:

One finalized graph about the Bonanza Creek snowshoe hare population (you pick which variables, how you want to wrangle it beforehand, and which type of visual to create - but make sure it is beautifully finalized)

One finalized HTML table (probably created using kable & kableExtra) containing summary statistics about the snowshoe hares (again, you pick which variables, and how you want to group/summarize them)

Make sure that both your figure and table appear in your final knitted document, each with a useful caption. Include text associated with each to help the audience understand and interpret the results. 


```{r}

library(tidyverse)
library(kableExtra)
library(janitor)
library(lubridate)

hares<- read_csv("showshoe_lter.csv") 

hares_mf <- hares %>% 
  filter(sex == "M" | sex == "F") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(full_date = date) %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(season = ifelse(month %in% c(1:4, 11, 12), "winter", "summer"))





```

```{r}

# graph

hares_plot <- ggplot(data = hares_mf, aes(x = year, y = weight))+
  geom_point(aes(color = sex))+
  facet_wrap(~season)

hares_plot

```







## Task 2. 

For this task, I’ll provide a couple of datasets that you *can* use for PCA exploration, but you are also welcome to find/choose a different dataset to use. You only need to use one dataset to practice running PCA, creating a biplot, and interpreting the results. Whichever dataset you choose, create a finalized HTML (knitted from .Rmd) that can be included as a project on your blogdown site including: 

A useful descriptive introductory summary (3 - 4 sentences) that helps the audience understand the data (include a citation as necessary) and what you’ll be exploring
All of your organized and well-annotated code (with warnings/messages hidden) used to wrangle then run PCA, then create a finalized PCA biplot that appears (with a caption) in your knitted HTML
A summary (bulletpoints are fine) pointing out some major takeaways from your PCA that can be gleaned from the biplot
You are also welcome to include a photo in the HTML to add some visual interest (not required for Task 2). 

